HERMES: Hybrid Error-corrector Model with inclusion of External Signals for nonstationary fashion time series

### 2023-04-24 tmlr_submission/hermes_paper_submission/hermes_paper.tex
add anonymized github link

\begin{abstract}
Developing models and algorithms to draw causal inference for time series is a long standing statistical problem. It is crucial for many applications, in particular for fashion or retail industries, to make optimal inventory decisions and avoid massive wastes. By tracking thousands of fashion trends on social media with state-of-the-art computer vision approaches, we propose a new model for fashion time series forecasting. Our contribution is  twofold. We first provide publicly\footnote[1]{\url{http://files.heuritech.com/raw_files/f1_fashion_dataset.tar.xz}} an appealing fashion dataset gathering \numberts\ weekly fashion time series. As influence dynamics are the key of emerging trend detection, we associate with each time series an external weak signal representing behaviours of influencers. Secondly, to leverage such a complex and rich dataset, we propose a new hybrid forecasting model\footnote[2]{\url{https://github.com/etidav/HERMES}}. Our approach combines per-time-series parametric models with seasonal components and a global recurrent neural network to include sporadic external signals. This hybrid model provides state-of-the-art results on the proposed fashion dataset, on the weekly time series of the M4 competition \citep{makridakis2018}, and illustrates the benefit of the contribution of external weak signals.
\end{abstract}

@@

With this paper, we provide publicly\footnote[1]{\url{http://files.heuritech.com/raw_files/f1_fashion_dataset.tar.xz}} a sample of $\numberts$ normalized fashion trends for men and women, over 9 different categories and 5 different markets. Each sequence has 261 time steps, from 2015-01-05 to 2019-12-31 with weekly values and no missing values. This collection of $\numberts$ fashion trends was selected in order to represent finely the issues faced by the fashion industry. For instance, some sequences show complex behaviours with sudden changes, referred to as emerging or declining trends. A central point of this work is to accurately detect and forecast such trends. In addition, each fashion time series is linked with its associated normalized fashion forward signal as presented in the section above. An overview of the dataset can be found in Table~\ref{tab:fashiondataset}.

@@

For the first parametric per-time-series models, existing Python  libraries named \texttt{statsmodels} and  \texttt{tbats} are used to estimate the different parameters $\statparam^n$.  Depending of the choice of local parametric models, two versions of HERMES are proposed. The first one uses as predictors an additive exponential smoothing model as a reference close to \citet{smyl2020}. The second one uses the TBATS model of \citet{alysha2011} and  achieves the highest accuracy results on the fashion dataset. The neural network architecture is composed of 3 LSTM layers of shape 50 and a final Dense layer to provide the correct output dimension. A classical Adam optimizer is used with a learning rate and a batch size set using a grid search. The loss function is defined as follows:
$$
\ell(\ts^n_{T+1:T+\lag},\tspred^n_{T+1:T+\lag|T}) = \frac{1}{\meants^n_T}\sum_{i=1}^{\lag}|\ts^n_{T+i} - \tspred^n_{T+i|T}|\,.
$$
This choice of $\mathrm{L}_1$ loss function is motivated by its robustness to outliers which accounts for some time series in the fashion industry with very specific behaviours. The loss and previous parameters are all set with a complete grid search. See \ref{sec:fashiongridsearch} for additional results concerning the loss function choice and \ref{sec:m4gridsearch} for a complete grid search example. The code is developed in Python using the Tensorflow library and publicly available\footnote[1]{\url{https://github.com/etidav/HERMES}}. It allows the use of GPU to speed up the training process.

>>>>

\begin{abstract}
Developing models and algorithms to draw causal inference for time series is a long standing statistical problem. It is crucial for many applications, in particular for fashion or retail industries, to make optimal inventory decisions and avoid massive wastes. By tracking thousands of fashion trends on social media with state-of-the-art computer vision approaches, we propose a new model for fashion time series forecasting. Our contribution is  twofold. We first provide publicly\footnote[1]{\url{https://anonymous.4open.science/r/HERMES-703F/}} an appealing fashion dataset gathering \numberts\ weekly fashion time series. As influence dynamics are the key of emerging trend detection, we associate with each time series an external weak signal representing behaviours of influencers. Secondly, to leverage such a complex and rich dataset, we propose a new hybrid forecasting model\footnotemark. Our approach combines per-time-series parametric models with seasonal components and a global recurrent neural network to include sporadic external signals. This hybrid model provides state-of-the-art results on the proposed fashion dataset, on the weekly time series of the M4 competition \citep{makridakis2018}, and illustrates the benefit of the contribution of external weak signals.
\end{abstract}

@@

With this paper, we provide publicly\footnote[1]{\url{https://anonymous.4open.science/r/HERMES-703F/}} a sample of $\numberts$ normalized fashion trends for men and women, over 9 different categories and 5 different markets. Each sequence has 261 time steps, from 2015-01-05 to 2019-12-31 with weekly values and no missing values. This collection of $\numberts$ fashion trends was selected in order to represent finely the issues faced by the fashion industry. For instance, some sequences show complex behaviours with sudden changes, referred to as emerging or declining trends. A central point of this work is to accurately detect and forecast such trends. In addition, each fashion time series is linked with its associated normalized fashion forward signal as presented in the section above. An overview of the dataset can be found in Table~\ref{tab:fashiondataset}.

@@

For the first parametric per-time-series models, existing Python  libraries named \texttt{statsmodels} and  \texttt{tbats} are used to estimate the different parameters $\statparam^n$.  Depending of the choice of local parametric models, two versions of HERMES are proposed. The first one uses as predictors an additive exponential smoothing model as a reference close to \citet{smyl2020}. The second one uses the TBATS model of \citet{alysha2011} and  achieves the highest accuracy results on the fashion dataset. The neural network architecture is composed of 3 LSTM layers of shape 50 and a final Dense layer to provide the correct output dimension. A classical Adam optimizer is used with a learning rate and a batch size set using a grid search. The loss function is defined as follows:
$$
\ell(\ts^n_{T+1:T+\lag},\tspred^n_{T+1:T+\lag|T}) = \frac{1}{\meants^n_T}\sum_{i=1}^{\lag}|\ts^n_{T+i} - \tspred^n_{T+i|T}|\,.
$$
This choice of $\mathrm{L}_1$ loss function is motivated by its robustness to outliers which accounts for some time series in the fashion industry with very specific behaviours. The loss and previous parameters are all set with a complete grid search. See \ref{sec:fashiongridsearch} for additional results concerning the loss function choice and \ref{sec:m4gridsearch} for a complete grid search example. The code is developed in Python using the Tensorflow library and publicly available\footnote[1]{\url{https://anonymous.4open.science/r/HERMES-703F/}}. It allows the use of GPU to speed up the training process.
